---
title: 'Assignment 13'
author: "Gracie Fischer"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment Details

### Purpose

The goal of this assignment is to practice working with geospatial
raster and vector data.

### Task

Write R code to successfully answer each question below.

### Criteria for Success

-   Code is within the provided code chunks or new code chunks are
    created where necessary
-   Code chunks run without errors
-   Code chunks have brief comments indicating which code is answering
    which part of the question
-   Code will be assessed as follows:
    -   Produces the correct answer using the requested approach: 100%
    -   Generally uses the right approach, but a minor mistake results
        in an incorrect answer: 90%
    -   Attempts to solve the problem and makes some progress using the
        core concept, but returns the wrong answer and does not
        demonstrate comfort with the core concept: 50%
    -   Answer demonstrates a lack of understanding of the core concept:
        0%
-   Any questions requiring written answers are answered with sufficient
    detail

### Due Date

**This Assignment is Optional!** April 22 at midnight MST

# Assignment Exercises

### 1. Working with raster data in terra (20 pts)

In this exercise, you will work with the same precipitation data as in
class, but this time focus in on the Himalayan region of Asia.

a.  Load in the terra package. What version does it say you are using?

    ```{r}
    library(terra)

    #R version 4.3.3
    ```

b.  Load in the precipitation data ("global_precipitation.tif"), what is
    the EPSG used in these data?

    ```{r}
    precip <- rast("global_precipitation.tif")

    precip

    #coord. ref. : lon/lat WGS 84 (EPSG:4326) 
    ```

c.  Define a geographic extent object for the Himalayan region.

    ```{r}
    himalaya_ext <- ext(c(68, 100, 25, 40))
    ```

d.  Crop the precipitation data to the Himalayan extent; what are
    resulting dimensions?

    ```{r}
    precip_himalaya <- crop(precip, himalaya_ext)

    precip_himalaya

    #dimensions  : 360, 768, 1  (nrow, ncol, nlyr)
    ```

### 2. Printing maps (20 points)

Still working with the precipitation data, this exercise focuses on
visualizing the data.

a.  Plot the precipitation data for the Himalayan region.

    ```{r}
    plot(precip_himalaya)
    ```

b.  Add a title to the plot.

    ```{r}
    plot(precip_himalaya, main = "Mean annual precipitation (mm)")
    ```

c.  Change the palette to that of `topo.colors` (use `?topo.colors` for
    details) and print the resultant plot.

    ```{r}
    plot(precip_himalaya, main = "Mean annual precipitation (mm)", col = topo.colors(n = 50))
    ```

d.  Change the palette to one from
    [ColorBrewer](https://colorbrewer2.org) for sequential (continuous)
    data.

    ```{r}
    plot(precip_himalaya, main = "Mean annual precipitation (mm)", col = hcl.colors(n = 50, palette = "GnBu"))
    ```

### 3. Working with vector data in terra (20 points)

a.  Create a data frame of three peaks in the Himalayan region: (lat,
    lon)

-   K2: 35.93, 76.50

-   Annapurna: 28.53, 83.81

-   Mount Everest: 28.05, 86.91

    ```{r}
    himalayan_peaks <- data.frame(
      peak = c("K2", "Annapurna", "Mount Everest"),
      lat = c(35.93, 28.53, 28.05),
      long = c(76.50, 83.81, 86.91)
    )

    head(himalayan_peaks)
    ```

b.  **Create a SpatVector object from the data frame with peak data.**

    ```{r}
    peaks_vector <- vect(himalayan_peaks, crs = crs(precip_himalaya))
    ```

c.  **Convert the points SpatVector object to a lines object.**

    ```{r}
    plot(precip_himalaya, col = temp_cols)
    plot(peaks_vector, add = TRUE)
    ```

d.  **Add the lines object to the precipitation plot of the Himalayan
    region.**

    ```{r}

    ```

### 4. Modifying raster values (20 points)

a.  Load in the elevation data ("global_elevation.tif"), what are the
    units for these data?

    ```{r}
    global_elev <- rast("global_elevation.tif")

    global_elev

    #Units are in meters
    ```

b.  Convert the units of the raster to feet.

    ```{r}
    global_elev_ft <- global_elev * 3.281
    ```

c.  Crop the elevation data to the same geographic extent as
    precipitation data.

    ```{r}
    elev_himalaya <- crop(global_elev_ft, himalaya_ext)
    ```

d.  Plot the Himalayan elevation data.

    ```{r}
    plot(elev_himalaya)
    ```

### 5. Converting raster objects to spatial vector objects (20 points)

For this last exercise, create a plot that shows precipitation for the
Himalayan region, using a color palette from ColorBrewer AND add a layer
that is partially (50%) transparent to shade out regions below 13,000
feet.

```{r}
plot(precip_himalaya, col = hcl.colors(n = 50, palette = "YlGnBu"),
     )

elevation_mask <- elev_himalaya < 13000
plot(elevation_mask, add = TRUE, col = rgb(0, 0, 0, alpha = 0.5))
```
